---
output:
  md_document:
    variant: gfm
---
### Using gen.simuHaplo_traceback

The traceback function allows the user to specify a proband and a founder, and given the output of `gen.simuHaplo`, the traceback function will go through all the simulations, identify the simulations in which the proband inherits any segment from the specified founder, and then identify the path of inheritance.  

For this example we will use the same "genea140" sample genealogy
```{r echo=TRUE, messages=FALSE, warning = FALSE, cache=TRUE}
library(GENLIB)
data(genea140)
gen_obj <- gen.genealogy(genea140)
gen_obj
```

For this example we will simulate a hypothetical chromosome 1, with a length of 290,000,000 BP and genetic length of 198cM for males, and 328cM for females. We will use a Poisson process as the model of meiosis, and we will produce the all_nodes output which is required for the traceback.
```{r message=FALSE, warning=FALSE, cache=TRUE}
gen.simuHaplo(gen_obj, model = 1, simulNo = 200, model_params = c(1.98, 3.28), 
              cM_len = c(198, 328), BP_len = 290000000, all_nodes = 1)
```

Now that we have the simulation results we can use the traceback function to investigate the relationship between any of the simulated probands and a select founder. For this illustration we arbitrarily pick a proband (can see a list of probands with `gen.pro()`), and one of their founders (can get founders with `gen.findFounders()`). For this example we will use proband 408992 and founder 863241. First let us visualize the relationship:

```{r, echo=TRUE, message=FALSE, warning=FALSE, cache=TRUE}
sub_genealogy <- gen.branching(gen_obj, 408992)
gen.graph(sub_genealogy, cex=0.5)
```

Now we can use the **gen.simuHaplo_traceback** function to identify every simulation where proband 408992 inherits a segment from founder 863241 and identify the path of inheritance. The function takes the following parameters:  
**gen: ** The genealogy object that was used for the simulation
**proID: ** The ID of the proband 
**ancestorID: ** ID of the founder. Any founder from this ancestor appearing in the proband haplotype will be traced back.
**all_nodes_path: ** The path to the 'All_nodes_haplotypes.txt' file created from the simulation.
**proband_haplotypes_path: ** The path to the 'Proband_Haplotypes.txt' file created from the simulation.

```{r, collapse=TRUE, echo=TRUE, cache=TRUE}
traceback <- gen.simuHaplo_traceback(gen_obj, proID =408992, ancestorID = 863241, 
                                     all_nodes_path = 'All_nodes_haplotypes.txt',
                                     proband_haplotypes_path = 'Proband_Haplotypes.txt')
```

The traceback function will print all the paths that it encounters in the simulation results, in the order that it encounters them. Note that if a path does not end with the ancestor ID that means the path bifurcates at the last internal ancestor that was printed. If an internal ancestor is homozygous for the founder segment it is possible for it to transmit a longer segment that has been stitched together through a recombination. In this case the traceback is terminated at this internal ancestor.
  
The function returns a dataframe that specifies the path, and the length of the segment.

```{r, echo=TRUE, cache=TRUE}
head(traceback)
```

